Quantifying the global film festival circuit: networks, diversity and value creation

### 2023-04-12 cudan/paperplots.qmd
values

```{r}
library(plotly)
library(tidyverse)
library(colorspace)
library(ggbeeswarm)
library(ggrepel)
library(patchwork)
library(visNetwork)
library(rworldmap)
library(scales)
library(igraph)
# library(lemon)

@@

abcols= divergingx_hcl(6, palette="Zissou 1")[c(1,5)] %>% c("gray2")
abshapes=c(16,4)

sixgcols = c(divergingx_hcl(6, palette="Zissou 1") %>% {.[6]=darken(.[6],0.3);.},  "gray50") %>%
  {.[1]=lighten(.[1],0.4);.} %>% 
  desaturate(0.07) %>% {.[c(2)]=darken(.[c(2)],0.1);.} 
sixcols = c(divergingx_hcl(6, palette="Zissou 1") %>% {.[c(1)]=darken(.[c(1)],0.3);.}, "gray10",  "gray50") 
langcols= c(divergingx_hcl(6, palette="Zissou 1")%>% {.[c(1)]=darken(.[c(1)],0.3);.}, "gray10",  "gray50")[c(1,3,5,6,7,8)] 
rolecols=divergingx_hcl(31, palette="Zissou 1")  %>% {c(.[1:2], "gray75", .[30:31])}

@@

@@

exgenfests = c("Festival de Cannes", "Sundance", "Hot Docs Festival", "Sitges FF")
exgeofests = c("Festival de Cannes", "Sundance", "Tokyo IFF", "Zurich FF")
exlangfests=c("Festival de Cannes", "Sundance", "Busan IFF", "BAFICI")

@@

div2plot = function(vecs, ex=NULL, convert=T,  colorlab="Main", xlab="External diversity", ylab="Internal diversity", genrecols=sixcols){
  # if(isgenre){
  #     vecs = vecs %>% arrange(YearEvent) %>% 
  #   mutate(colorgenre = 
  #          case_when(firstkind %in% head(names(sort(table(.$firstkind),decreasing = T)),6)~firstkind, T~"other" ) %>% as.factor()) %>% 
  #   mutate(colorgenre = fct_relevel(colorgenre, 
  #                                 c(names(sort(table(colorgenre),decreasing = T)) %>% .[.!="other"], "other") ))
  #     colorlab="Main\ngenre"
  # } else {
  #   vecs = vecs %>% arrange(YearEvent) %>% 
  #   mutate(colorgenre = 
  #          case_when(firstcountry %in% head(names(sort(table(.$firstcountry),decreasing = T)),6)~firstcountry, T~"other" ) %>% as.factor()) %>% 
  #   mutate(colorgenre = fct_relevel(colorgenre, 
  #                                 c(names(sort(table(colorgenre),decreasing = T)) %>% .[.!="other"], "other") ))

@@

  g={ggplot(vecs , 
        aes(exdiv, indiv, shape=ab, color=colorgenre, size=ab, text=tool))+
      geom_vline(xintercept = median(vecs$exdiv, na.rm=T) # mean(c(min(vecs$exdiv,na.rm=T), max(vecs$exdiv,na.rm=T)))
                 , color="gray25")+
      geom_hline(yintercept = median(vecs$indiv, na.rm=T)
                #   mean(c(min(vecs$indiv,na.rm=T), max(vecs$indiv,na.rm=T)))
                 , color="gray25")+
      geom_point(#data=vecs %>% filter(!(libelleFestival_NEW_VZ%in%ex)), 
                 alpha=0.7) +
      #geom_point(data=vecs %>% filter((libelleFestival_NEW_VZ%in%ex)), alpha=1)+
    #geom_line(aes(YearEvent, indiv), data=festvecs %>% group_by(YearEvent) %>% summarise(indiv=mean(indiv,na.rm=T)), inherit.aes = F)+
    # geom_path(aes(group=libelleFestival_NEW_VZ), 
    #           vecs %>% filter(libelleFestival_NEW_VZ%in%ex) ,
    #           color="black", 
    #           size=0.2)+
    #scale_color_manual(values=abcols)+
    scale_color_manual(values=genrecols, name=colorlab)+
    scale_shape_manual(values=abshapes, name="Festival\ntype")+
    scale_size_manual(values=c(2, 1.2), guide = "none")+
    labs(x=xlab, y=ylab)+
    theme_bw()+
    theme(#legend.position = 'left', 
      panel.border = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.background = element_rect(color="white", fill="white")
    )} 
  if(convert){
    ggplotly(g,tooltip="text")
  } else {
    g
  }
}

divyearplot = function(vecs, ex, shape=16, yl="", splitab=T, convert=T, networkstats=T, ylims=NULL){
  if(splitab){
    vecs = vecs %>% mutate(YearEvent=case_when(ab=="A"~YearEvent-0.15,
                                               ab=="B"~YearEvent+0.15 )) %>% 
      arrange(libelleFestival_NEW_VZ) %>% 
      mutate(series = substr(libelleFestival_NEW_VZ, 1,15))
    if(length(unique(vecs$libelleFestival_NEW_VZ)) != 
       length(unique(vecs$series))){ stop("shortening mismatch")}
    g=ggplot(vecs , 
        aes(YearEvent, target, color=ab, shape=ab,  text=tool))
  } else {
     vex = vecs %>% 
       filter(libelleFestival_NEW_VZ %in% ex) %>% 
       mutate(libelleFestival_NEW_VZ=fct_relevel(libelleFestival_NEW_VZ, ex))
     g=ggplot(vex , 
        aes(YearEvent, target, color=ab, shape=ab,  text=tool))
  }

@@

      panel.grid.major.x  = element_blank(),
       panel.grid.minor.x  = element_line(color="gray80"),
      plot.background = element_rect(color="white", fill="white"),
      axis.title.x = element_blank()
    )+
    labs(y=yl)
  # if(nrow( vecs %>% filter(libelleFestival_NEW_VZ%in%ex))>0){
  #   g=g+geom_line(aes(group=libelleFestival_NEW_VZ), 
  #             vecs %>% filter(libelleFestival_NEW_VZ%in%ex) ,
  #             color="gray40", 
  #             size=0.2)
  # }

  if(!splitab & !is.null(ex)){
   if(networkstats){
      fromy=2014.5; toy=2019.5
    } else{
      fromy=2008; toy=2022
      g=g+geom_errorbar(aes(ymin=lower, ymax=upper), color="gray40", width=0.1, size=0.45)
    }
    g=
      g+facet_wrap(~libelleFestival_NEW_VZ, nrow = ceiling(length(ex)/6))+
      # geom_point( aes(YearEvent, target),data=vecs %>% select(-libelleFestival_NEW_VZ), inherit.aes = F,
      #             color="gray90", size=0.9)+
      geom_line(data=vex %>% filter(YearEvent>fromy, YearEvent<toy) %>% 
                  group_by(YearEvent, libelleFestival_NEW_VZ) %>% 
                  summarise(target=mean(target, na.rm=T), tool=NA, ab=ab[1]), 
                color="gray60", size=0.3
                )+
      geom_point(data=vex, alpha=1, size=1.5, shape=16)+
      scale_x_continuous(breaks=seq(2009,2021,3) )+
      scale_y_continuous(limits=c( min(vecs$target), max(vecs$target) ))+
      theme(
       panel.grid.major.x  = element_line(color="gray95"),
       panel.grid.minor.x  = element_blank(),
       legend.position = 'none'
       )
  }

  if(!is.null(ylims)){
    g=g+scale_y_continuous(limits = ylims)
  }

@@

  g=  ggplot(sysdiv,aes(YearEvent, sysdiv, color=ab))+
    #geom_pointline(position=position_dodge(width = 0.25), distance=unit(10, "pt") )+
    geom_line(data=sysdiv %>% filter(ab=="Both"), position=position_nudge(x = 0.25/3))+
    geom_point(position=position_dodge(width = 0.25), size=0.5)+
    geom_errorbar(aes(ymin=slower, ymax=supper),width=0.1, linewidth=0.45,position=position_dodge(width = 0.25) )+
    scale_x_continuous(breaks=2012:2021)+
    scale_color_manual(values=c(abcols) %>% {.[1:2]=lighten(.[1:2],0.2);. })+
    theme_bw()+
    theme(legend.position = 'none')+
    theme(
      panel.grid.minor.x  = element_blank(),
       panel.grid.minor.y  = element_blank(),
      plot.background = element_rect(color="white", fill="white"),
      axis.title.x = element_blank()
    )+
    labs(y=yl)
 return(g)
}

gg=function(g){
  ggplotly(g, tooltip="text") %>%  plotly::config(doubleClickDelay=500)
}

@@

# big force directed, use that.
gr = as_tbl_graph(graph_from_data_frame(e2, directed = F, vertices = nodelist %>% filter(EventID_VZ %in% c(e2$from,e2$to))))
g = ggraph(gr  ) +
  geom_node_point(aes(color=YearEvent, size=nf)) +
  #geom_node_text(aes(label=name)) +
  geom_edge_arc(aes(color=coloryear, alpha=dist),  width=0.1, strength=0.04) +
  scale_color_viridis(end=0.9)+
  scale_edge_color_viridis(end=0.9)+
  scale_edge_alpha(range=c(0.15, 0.5))+
  scale_size(0.5, 1.5)+
  theme_void()+
  theme(legend.position = 'none',
        plot.background = element_rect(color="white", fill="white")
  )

ggsave("yearnetwork2.png", g, width=5000, height=5000, scale=0.6, units="px")
# write_csv(e2, "edgelist_nov22.csv")

@@

Color is n shared films (edge strength). This is actually more informative than a network plot, as it easily shows both degree and strength for all pairs of festivals/nodes, while a network plot

-   hides information due to overlaps, such as long-distance links
-   can only effectively show degree (strength as edge weight is lost is the mess)
-   can create incidental visual artefacts that can mislead interpretation due to the semi-random layout.

So tbh we should probably use a matrix instead, but I'm aware networks look aesthetically cooler, and that some of us have interest in having a network-person profile. So maybe a network and a matrix side by side, would that be useful...? Interactive networks are of course a bit better, as you can zoom into the mess (try above).

```{r, fig.width=7, fig.width=7}
hm=heatmaplist %>% 
  mutate(fillweight=log10(weight)) %>% 
         mutate(tool=paste0(tolibelle, "\n", fromlibelle, "\nn=",weight, " dist=",dist ))
hmy = hm %>% 
        arrange(tolibelle) %>% 
          group_by(toyear) %>% 
          filter(1:n()==round(n()/2)) %>% 
  mutate(fromlibelle=tolibelle)
cls = sixcols[c(1,1,2,3)] %>% {.[1]=darken(.[1],0.6);.} %>% 
                         {.[4]=lighten(.[4],0.5);.} %>% 
  rev

{ggplot(hm , 
       aes(fromlibelle, tolibelle, fill=weight, text=tool))+
  coord_cartesian(expand=F)+
  geom_tile(size=0)+
  geom_text(aes(label=toyear, x=0), data=hmy, size=2, hjust=0,vjust=1, angle=90)+
  geom_text(aes(label=toyear,x=fromlibelle, y=0), data=hmy, size=2, vjust=0)+
  #scale_fill_viridis_c(end=0.9, na.value = "white")+
    scale_fill_gradientn(colors = cls, na.value = "white", trans="log10", name="log\nstrength")+
  theme_bw()+
    theme(legend.position = 'none', 
      panel.border = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.background = element_rect(color="white", fill="white"),
      axis.title = element_blank(),
      axis.text = element_blank(), # element_text(size=3),
      axis.ticks = element_blank()
     # axis.text.x = element_text(angle=90, hjust=1, vjust=1)
    )} %>% ggplotly(tooltip = "text")

# static
ggplot(hm , 
       aes(fromlibelle, tolibelle, fill=weight, text=tool))+
  coord_cartesian(expand=F)+
  geom_tile(size=0)+
  geom_text(aes(label=toyear, x=0), data=hmy, hjust=0,vjust=1, angle=90,size=3, 
            color=viridis_pal(end = 0.9)(13))+
  geom_text(aes(label=toyear,x=fromlibelle, y=0), data=hmy,vjust=0, size=3, 
            color=viridis_pal(end = 0.9)(13))+
  #scale_fill_viridis_c(end=0.9, na.value = "white")+
  #scale_fill_gradientn(colors = cls, na.value = "white", trans="log10", name="log\nstrength")+
  scale_fill_viridis_c(direction=-1,na.value = "white", option="magma", trans="log10", name="log\nstrength")+
  theme_bw()+
    theme(legend.position = c(0.9,0.3), 
      panel.border = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.background = element_rect(color="white", fill="white"),
      axis.title = element_blank(),
      axis.text = element_blank(), # element_text(size=3),
      axis.ticks = element_blank()
    )

@@

Interactive nodes yearwise, with hover labels:

```{r}
gmp=ggplot(nodelist, aes(x, y,color=YearEvent,text=tool, shape=ab, size=ab))+
  geom_point( alpha=0.8) + 
  scale_x_continuous(breaks=2009:2021)+
  scale_color_viridis_c(end=0.9)+
  scale_shape_manual(values=abshapes)+
  scale_size_manual( values=c(2, 1))+
  theme_void()+
  theme(legend.position = 'none', axis.text.x=element_text(),
        plot.background = element_rect(color="white", fill="white")
  )
plotly::ggplotly(gmp, tooltip="text")

@@

Probably not that useful:

@@

# Network stats

@@

Without any links, degree=0:
```{r}
nodelist %>% filter(degree==0)
```

@@

# Content types ("genres" in the database)

@@

g=ggplot( mapping=aes(V1, V2, color=colorgenre))+
  geom_point(aes(shape=ab, text=tool), data=ugenre %>% filter(type=="festival"), size=2, fill=NA, alpha=0.6)+
  # geom_shadowtext(aes(label=firstkind, size=sizegen), 
  #                 data=ugenre %>% filter(type %in% c("genre")), 
  #                 bg.color="white", alpha=0.5, hjust=0, show.legend = FALSE
  #             # ,max.overlaps=100, box.padding = 0, point.padding = 0, min.segment.length = 999, force_pull=0.5, max.time = 2
  #             )+
  geom_point( data=ugenre %>% filter(type=="film"), size=0.1, fill=NA,show.legend = FALSE)+
  geom_text_repel(aes(label=event, size=sizegen %>% log), 
                  data=ugenre %>% filter(type %in% c("genre", "film")), 
            lineheight=0.49,
            alpha=1, hjust=0, 
            #bg.color="white",  
            show.legend = FALSE, max.overlaps=100, box.padding = 0, point.padding = 0, min.segment.length = 999, force_pull=0.5, max.time = 2)+

@@

g

gg(g)

@@

Possible alternative: if drama is main, add blue ring around point, but show secondary as main color.

@@

# festival illustration squares.
ggplot(tops, aes( x=s,y=libelleFestival_NEW_VZ, fill=colorgenre2))+
  geom_bar(stat="identity", color=NA,position="stack")+
  scale_fill_manual(values=
                      c(sixgcols %>% head(-1),
                        gray.colors(
                          length(unique(tops$colorgenre2))-
                            length(mains),
                          start=0.5, end=0.7
                          )) %>% rev)+
  coord_cartesian(expand = 0, xlim=c(0,100))+
  theme_void()+
  theme(axis.title = element_blank(),
        legend.position = "none",
        panel.border  = element_rect(size=0.5, color="gray10", fill=NA)
        )+
  facet_wrap(~libelleFestival_NEW_VZ, strip.position="bottom", scales = "free", nrow=2)

@@

@@

divyearplot(festgenrevecs  %>% mutate(target=indiv),
            ex=NULL, yl="Internal genre diversity", convert=convertplots, networkstats = F)
divyearplot(festgenrevecs  %>% mutate(target=indiv, lower=ilower, upper=iupper), 
            ex=exgenfests, yl="Internal genre diversity", convert=T, splitab=F, networkstats = F)

@@

>>>>

```{r}
library(plotly)
library(tidyverse)
library(colorspace)
library(ggbeeswarm)
library(ggrepel)
library(patchwork)
library(visNetwork)
library(rworldmap)
library(scales)
library(igraph)
library(ggraph)
library(tidygraph)
library(PieGlyph)
# library(lemon)

@@

@@

abcols= divergingx_hcl(6, palette="Zissou 1")[c(1,4)] %>% c("gray2") %>% desaturate(0.6) #%>% show_col
abshapes=c(16,4)

sixgcols = c(divergingx_hcl(6, palette="Zissou 1") %>% {.[6]=darken(.[6],0.3);.},  "gray80") %>%
  {.[1]=lighten(.[1],0.2);.} %>% {.[3] = "#03408f";.} %>% 
  {.[c(2)]=darken(.[c(2)],0.1);.} %>% 
  {.[c(1,4,2,5,3,6,7)]} # %>%  show_col
sixcols = c(divergingx_hcl(6, palette="Zissou 1"), "gray10",  "gray80") %>% {.[c(1)]=lighten(.[c(1)],0.5);.} %>% {.[c(2)]=darken(.[c(1)],0.5);.} %>% {.[c(4)]=lighten(.[c(4)],0.2);.} %>% {.[7]=darken(.[1],0.75);.} # %>% show_col
langcols= c(divergingx_hcl(6, palette="Zissou 1") %>% 
              {.[c(1)]=darken(.[c(1)],0);.} %>% 
              {.[7]=darken(.[1],0.6);.} ,  "gray80")[c(1,3,5,6,7,8)] #%>% show_col
rolecols=divergingx_hcl(31, palette="Zissou 1")  %>% {c(.[1:2], "gray75", .[30:31])}
matcols = (divergingx_hcl(8, palette="Zissou 1")[c(3,5,7)] %>% {.[1]=lighten(.[1],0.4);.} %>%  {.[3]=darken(.[3],0.95);.}) # %>% show_col
netcols =(divergingx_hcl(5, palette="Zissou 1")[c(1,1,2,3)] %>% {.[1]=darken(.[1],0.5);.} %>%  {.[4]=lighten(.[4],0.45);.} ) %>% {c("#01033a", .)} #  %>% show_col

@@

exgenfests = c("Festival de Cannes", "Sundance", "Hot Docs Festival", "Sitges FF")
exgeofests = c("Festival de Cannes", "Sundance", "Tokyo IFF", "Zurich FF")
exlangfests=c("Festival de Cannes", "Sundance", "Busan IFF", "BAFICI")
exrolefests = c("Festival de Cannes", "Sundance","SXSW",  "Berlinale - Berlin IFF")

# gender examples TODO------- aside from Cannes and Sundance (and instead of Hot Docs and Sitges), could we please have SXSW for B and Berlinale for A

@@

### functions ####

dopies = function(vecs, tops=NULL){
  mains=levels(vecs$colorgenre) 
  if( !attr(vecs, "tops") %>% is.null){
    tops = attr(vecs, "tops") 
  }
  pies = tops %>% 
   left_join(vecs %>% filter(!duplicated(EventID_VZ)), by="EventID_VZ") %>% 
   select(txtKind, s, EventID_VZ, libelleFestival_NEW_VZ, YearEvent, LibelleEvent_NEW_VZ)   %>% 
  mutate(txtKind=case_when(txtKind%in%mains~txtKind, T~"other")) %>% 
  group_by(EventID_VZ, txtKind) %>% 
  summarise(s=sum(s), 
            across(c(LibelleEvent_NEW_VZ, libelleFestival_NEW_VZ, YearEvent), first), )    %>% 
  ungroup() %>% 
  mutate(colorgenre2=fct_relevel(txtKind, mains)) %>% select(-txtKind) %>% 
  arrange(colorgenre2)

  pies2 = pivot_wider(pies , names_from = colorgenre2, values_from=s, values_fill=0 ) %>%
  mutate(event=LibelleEvent_NEW_VZ) %>% 
  group_by(libelleFestival_NEW_VZ) %>% 
  arrange(YearEvent, .by_group = T) %>% 
  ungroup
  attr(pies2, "mains")=mains
return(pies2)
}

div2plot = function(vecs, ex=NULL, convert=T,  colorlab="Main", xlab="Contributing diversity", ylab="Internal diversity", genrecols=sixcols, pies){
  # if(isgenre){
  #     vecs = vecs %>% arrange(YearEvent) %>% 
  #   mutate(colorgenre = 
  #          case_when(firstkind %in% head(names(sort(table(.$firstkind),decreasing = T)),6)~firstkind, T~"other" ) %>% as.factor()) %>% 
  #   mutate(colorgenre = fct_relevel(colorgenre, 
  #                                 c(names(sort(table(colorgenre),decreasing = T)) %>% .[.!="other"], "other") ))
  #     colorlab="Main\ngenre"
  # } else {
  #   vecs = vecs %>% arrange(YearEvent) %>% 
  #   mutate(colorgenre = 
  #          case_when(firstcountry %in% head(names(sort(table(.$firstcountry),decreasing = T)),6)~firstcountry, T~"other" ) %>% as.factor()) %>% 
  #   mutate(colorgenre = fct_relevel(colorgenre, 
  #                                 c(names(sort(table(colorgenre),decreasing = T)) %>% .[.!="other"], "other") ))

@@

  g={ggplot(vecs , 
        aes(exdiv, indiv, shape=ab, color=colorgenre, size=ab, text=tool))+
      geom_vline(xintercept = median(vecs$exdiv, na.rm=T) # mean(c(min(vecs$exdiv,na.rm=T), max(vecs$exdiv,na.rm=T)))
                 , color="gray25")+
      geom_hline(yintercept = median(vecs$indiv, na.rm=T)
                #   mean(c(min(vecs$indiv,na.rm=T), max(vecs$indiv,na.rm=T)))
                 , color="gray25")+
      geom_point(#data=vecs %>% filter(!(libelleFestival_NEW_VZ%in%ex)), 
                 alpha=0.7) +
      #geom_point(data=vecs %>% filter((libelleFestival_NEW_VZ%in%ex)), alpha=1)+
    #geom_line(aes(YearEvent, indiv), data=festvecs %>% group_by(YearEvent) %>% summarise(indiv=mean(indiv,na.rm=T)), inherit.aes = F)+
    # geom_path(aes(group=libelleFestival_NEW_VZ), 
    #           vecs %>% filter(libelleFestival_NEW_VZ%in%ex) ,
    #           color="black", 
    #           size=0.2)+
    #scale_color_manual(values=abcols)+
    scale_color_manual(values=genrecols, name=colorlab)+
    scale_shape_manual(values=abshapes, name="Festival")+
    scale_size_manual(values=c(2, 1.2), guide = "none")+
    labs(x=xlab, y=ylab)+
    theme_bw()+
    theme(#legend.position = 'left', 
      #panel.border = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.background = element_rect(color="white", fill="white")
    )} 
  if(convert){
    ggplotly(g,tooltip="text")
  } else {
    g
  }
}

divyearplot = function(vecs, ex, shape=16, yl="", splitab=T, convert=T, networkstats=T, ylims=NULL, genrecols=sixgcols, pies=NULL, nrowmod=6){
  if(splitab){
    vecs = vecs %>% mutate(YearEvent=case_when(ab=="A"~YearEvent-0.15,
                                               ab=="B"~YearEvent+0.15 )) %>% 
      arrange(libelleFestival_NEW_VZ) %>% 
      mutate(series = substr(libelleFestival_NEW_VZ, 1,15))
    if(length(unique(vecs$libelleFestival_NEW_VZ)) != 
       length(unique(vecs$series))){ stop("shortening mismatch")}
    vex=vecs
    g=ggplot(vecs , 
        aes(YearEvent, target, color=ab, shape=ab,  text=tool))
  } else {
     vex = vecs %>% 
       filter(libelleFestival_NEW_VZ %in% ex, !is.na(target)) %>% 
       mutate(libelleFestival_NEW_VZ=fct_relevel(libelleFestival_NEW_VZ, ex))
     g=ggplot(vex , 
        aes(YearEvent, target, color=ab, shape=ab,  text=tool))
  }

@@

      panel.grid.major.x  = element_blank(),
       panel.grid.minor.x  = element_line(color="gray80"),
      plot.background = element_rect(color="white", fill="white"),
      axis.title.x = element_blank(),
      axis.ticks.length = unit(0, "inch")
    )+
    labs(y=yl)
  # if(nrow( vecs %>% filter(libelleFestival_NEW_VZ%in%ex))>0){
  #   g=g+geom_line(aes(group=libelleFestival_NEW_VZ), 
  #             vecs %>% filter(libelleFestival_NEW_VZ%in%ex) ,
  #             color="gray40", 
  #             size=0.2)
  # }

  if(!splitab & !is.null(ex)){
   if(networkstats){
      fromy=2014.5; toy=2019.5
    } else{
      fromy=2008; toy=2022
      g=g+geom_errorbar(aes(ymin=lower, ymax=upper), color="gray40", width=0.1, size=0.45)
    }
    g=
      g+facet_wrap(~libelleFestival_NEW_VZ, nrow = ceiling(length(ex)/nrowmod))+
      # geom_point( aes(YearEvent, target),data=vecs %>% select(-libelleFestival_NEW_VZ), inherit.aes = F,
      #             color="gray90", size=0.9)+
      geom_line(data=vex %>% filter(YearEvent>fromy, YearEvent<toy) %>% 
                  group_by(YearEvent, libelleFestival_NEW_VZ) %>% 
                  summarise(target=mean(target, na.rm=T), tool=NA, ab=ab[1]), 
                color="gray60", size=0.3
                )+
      geom_point(data=vex, alpha=1, size=1.5, shape=16)+
      theme(
       panel.grid.major.x  = element_line(color="gray95"),
       panel.grid.minor.x  = element_blank(),
       panel.grid.minor.y  = element_blank(),
       legend.position = 'none',
       strip.text = element_text(hjust=0, margin=margin(0.5,0,0.5,1))
       )

    if(!is.null(pies)){
      p = dopies(vex, pies)
      m = attr(p, "mains")
      vp = vex %>%  select(EventID_VZ, target,YearEvent,libelleFestival_NEW_VZ) %>% 
        left_join(p %>% select(EventID_VZ, {{m}}),by="EventID_VZ")
      g = g+ geom_pie_glyph(aes(YearEvent, target), data = vp, slices=m, radius=0.2, inherit.aes = F)+
        scale_fill_manual(values=genrecols, breaks=m)

    }
  }

  minx = min(vex$YearEvent, na.rm=T)
  if(!is.null(ylims)){
    g=g+
      coord_cartesian(ylim = ylims, expand=F)+
      scale_x_continuous(breaks=seq(minx+1,2020,3), limits=c(minx-0.4,2021+0.4 ) )
  } else {
    g=g+ scale_x_continuous(breaks=seq(minx+1,2020,3), expand = expansion(0.1,0) )+
      scale_y_continuous(expand=c(0.05,0)
        #limits=c( min(c(vecs$target, vecs$lower), na.rm=T), max(c(vecs$target, vecs$upper), na.rm=T )*1.001 )
        )
  }

@@

  g=  ggplot(sysdiv,aes(YearEvent, sysdiv, color=ab))+
    #geom_pointline(position=position_dodge(width = 0.25), distance=unit(10, "pt") )+
    geom_line(data=sysdiv %>% filter(ab=="Both"), position=position_nudge(x = 0.25/3))+
    geom_point(aes(shape=ab, size=ab), position=position_dodge(width = 0.25))+
    geom_errorbar(aes(ymin=slower, ymax=supper),width=0.1, linewidth=0.45,position=position_dodge(width = 0.25) )+
    scale_x_continuous(breaks=2012:2021)+
    scale_color_manual(values=abcols)+
    scale_shape_manual(values=abshapes %>% c(1))+
    scale_size_manual(values=c(0.6, 0.3, 0.7))+
    theme_bw()+
    theme(legend.position = 'none')+
    theme(
      panel.grid.minor.x  = element_blank(),
       panel.grid.minor.y  = element_blank(),
      plot.background = element_rect(color="white", fill="white"),
      axis.title.x = element_blank()
    )+
    labs(y=yl)
 return(g)
}

gg=function(g, do=T){
  if(do){
    ggplotly(g, tooltip="text") %>%  plotly::config(doubleClickDelay=500)
  } else{
    g
  }
}

pspacer=ggplot()+theme_void()+theme(plot.margin = margin(0,0,0,0))

@@

@@

Color is overlap between festival programmes, each row is a festival, cells show how much it (relative to itself) overlaps with past and future festivals.

```{r, fig.width=7, fig.width=7}
hm=heatmaplist %>% 
  #group_by(fromyear) %>%  arrange(-overlap, .by_group = T) %>%  mutate(y = as.factor((100*fromyear)+(1:n())) ) %>% 
  mutate(fillweight=log10(weight)) %>% 
  mutate(overlap2 = case_when(overlap==0 ~ NA, T~overlap)) %>% 
         mutate(tool=paste0(fromlibelle,"\n",  
                            tolibelle, 
                            "\nn shared=",weight,
                            "\n% overlap=",overlap, "\ntime dist=",dist ))

hmy = hm %>% # year labels
        arrange(tolibelle) %>% 
          filter(toyear>2011, fromyear>2011) %>% 
          group_by(toyear ) %>% 
          filter(1:n()==round(n()/2)) %>% 
  mutate(fromlibelle=tolibelle)

# {ggplot(hm , 
#        aes(tolibelle, fromlibelle, fill=overlap, text=tool))+
#   coord_cartesian(expand=F)+
#   geom_tile(size=0)+
#   geom_text(aes(label=toyear, x=0), data=hmy, size=2, hjust=0,vjust=1, angle=90)+
#   geom_text(aes(label=toyear,x=fromlibelle, y=0), data=hmy, size=2, vjust=0)+
#   #scale_fill_viridis_c(end=0.9, na.value = "white")+
#   #scale_fill_gradientn(colors = cls, na.value = "white", trans="log10", name="log\nstrength")+
#   scale_fill_gradientn(colors = c("white",cls), na.value = "white",  name="overlap")+
#   theme_bw()+
#     theme(legend.position = 'none', 
#       panel.border = element_blank(),
#       panel.grid.minor.x = element_blank(),
#       plot.background = element_rect(color="white", fill="white"),
#       axis.title = element_blank(),
#       axis.text = element_blank(), # element_text(size=3),
#       axis.ticks = element_blank()
#      # axis.text.x = element_text(angle=90, hjust=1, vjust=1)
#     )} %>% ggplotly(tooltip = "text")

fig_heatmap =  ggplot(hm %>% mutate(w=rescale(overlap2,to=c(1,1.3))) , 
       aes(tolibelle, fromlibelle, 
         fill=overlap2, text=tool))+
  coord_cartesian(expand=F)+
  geom_tile(aes(height=w, width=w), size=0)+
  geom_text(aes(label=toyear, x=0), data=hmy, hjust=0,vjust=1, angle=90,size=2.5)+ 
            #color=viridis_pal(end = 0.9)(13))+
  geom_text(aes(label=toyear,x=fromlibelle, y=0), data=hmy,vjust=0, size=2.5)+

  #scale_fill_viridis_c(end=0.9, na.value = "white")+
  #scale_fill_gradientn(colors = cls, na.value = "white", trans="log10", name="log\nstrength")+
  # scale_fill_viridis_c(direction=-1,na.value = "white", option="magma", trans="log10", name="log\nstrength")+
  scale_fill_gradientn(colors = matcols, na.value = "transparent",  name="Overlap %")+
  annotate("text", -Inf, Inf, hjust=-2, vjust=1, label="C", fontface='bold')+
  theme_bw()+
    theme(legend.position = c(0.94,0.22), 
      panel.border = element_blank(),
      panel.grid = element_blank(),
      plot.background = element_rect(color="white", fill="white"),
      axis.title = element_blank(),
      axis.text = element_blank(), # element_text(size=3),
      axis.ticks = element_blank()
    )

gg(fig_heatmap, convertplots)

@@

```{r, eval=F}
# Interactive nodes yearwise, with hover labels:
gmp=ggplot(nodelist, aes(x, y,color=YearEvent,text=tool, shape=ab, size=ab))+
  geom_point( alpha=0.8) + 
  scale_x_continuous(breaks=2009:2021)+
  scale_color_viridis_c(end=0.9)+
  scale_shape_manual(values=abshapes)+
  scale_size_manual( values=c(2, 1))+
  theme_void()+
  theme(legend.position = 'none', axis.text.x=element_text(),
        plot.background = element_rect(color="white", fill="white")
  )
plotly::ggplotly(gmp, tooltip="text")

@@

Network laid on on the world map according to event locations:

@@

# Network statistics

@@

```{r, eval=F}
# Without any links, degree=0:
nodelist %>% filter(degree==0)
```

@@

# Thematic or content types

@@

ggplot( mapping=aes(V1, V2, color=colorgenre))+
  geom_point(aes(shape=ab, text=tool), data=ugenre %>% filter(type=="festival"), size=2, fill=NA, alpha=0.6)+
  # geom_shadowtext(aes(label=firstkind, size=sizegen), 
  #                 data=ugenre %>% filter(type %in% c("genre")), 
  #                 bg.color="white", alpha=0.5, hjust=0, show.legend = FALSE
  #             # ,max.overlaps=100, box.padding = 0, point.padding = 0, min.segment.length = 999, force_pull=0.5, max.time = 2
  #             )+
  geom_point( data=ugenre %>% filter(type=="film"), size=0.1, fill=NA,show.legend = FALSE)+
  geom_text_repel(aes(label=event, size=sizegen %>% log), 
                  data=ugenre %>% filter(type %in% c("genre", "film")), 
            lineheight=0.49,
            alpha=1, hjust=0, 
            #bg.color="white",  
            show.legend = FALSE, max.overlaps=100, box.padding = 0, point.padding = 0, min.segment.length = 999, force_pull=0.5, max.time = 2)+

@@

gg(g, convertplots)

@@

@@

# festival illustration squares.
ggplot(tops, aes( x=s,y=libelleFestival_NEW_VZ, fill=colorgenre2))+
  geom_bar(stat="identity", color=NA,position="stack")+
  scale_fill_manual(values=
                      c(sixgcols %>% head(-1),
                        gray.colors(
                          length(unique(tops$colorgenre2))-
                            length(mains),
                          start=0.5, end=0.7
                          )) %>% rev)+
  coord_cartesian(expand = 0, xlim=c(0,100))+
  theme_void()+
  theme(axis.title = element_blank(),
        legend.position = "none",
        panel.border  = element_rect(linewidth=0.5, color="gray10", fill=NA)
        )+
  facet_wrap(~libelleFestival_NEW_VZ, strip.position="bottom", scales = "free", nrow=2)

@@

## Diversity

@@

divyearplot(festgenrevecs  %>% mutate(target=indiv),
            ex=NULL, yl="Internal genre diversity", convert=convertplots, networkstats = F)
divyearplot(festgenrevecs  %>% mutate(target=indiv, lower=ilower, upper=iupper), 
            ex=exgenfests, yl="Internal genre diversity", convert=convertplots, splitab=F, networkstats = F)

@@

## Distance between event and film production

Events in Europe, colored by average distance between event location and the production location of the programmed events

```{r}
# fc = festcountryvecs %>% filter(eventlong > -10, eventlong < 30, eventlat > 35, eventlat < 60)
# 
# fcvals = fc %>% mutate(txtCountryEvent_VZ=case_when(txtCountryEvent_VZ=="uk"~"united kingdom", T~txtCountryEvent_VZ)) %>% group_by(txtCountryEvent_VZ) %>% summarise(val=mean(eventdist,na.rm=T)) %>% ungroup %>% rename(id=txtCountryEvent_VZ) %>% select(id, val)
# mp2 = mp %>% mutate(id=tolower(id)) %>% left_join(fcvals, by="id")

# attempt to scatter, but still all on capitals, not great.
  # mutate(gr = paste(eventlong, eventlat)) %>% group_by(gr) %>%
  # mutate(ng = log2(n())) %>% 
  # mutate(eventlong2=case_when(
  #   n()<= 3 ~eventlong, 
  #   T~eventlong+
  #     (rep( seq(-floor(ng[1])/2, ceiling(ng[1])/2, length.out=ceiling(ng[1])), each=round(ng[1]) )[1:n()]/3 )
  #   ) ) %>% 
  # mutate(eventlong2=case_when(is.na(eventlong2)~eventlong, T~eventlong2))
